---
title: "Covid cases and deaths for selected countries and months"
author: "Marc A.T. Teunis, PhD"
date: "`r Sys.time()`"
output: html_document
runtime: shiny
params:
  country_code: NLD BEL
  from_to_date: 2020-01-01 2021-04-01
resource_files:
- app/shiny/app.R
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  error = FALSE
  
  )

```

## Packages
```{r}
library(utils)
library(tidyverse)
library(tools)
library(glue)
library(blscrapeR)
library(aweek)
library(lubridate)
```

## Data
Download data from European Center for Disease Prevention and Control [ECDC](https://www.ecdc.europa.eu/en)

[The link to the data file](https://opendata.ecdc.europa.eu/covid19/nationalcasedeath/csv)

[Data Dictionary](https://www.ecdc.europa.eu/sites/default/files/documents/2021-01-13_Variable_Dictionary_and_Disclaimer_national_weekly_data.pdf) 

See also the file: `./data-raw/D030/supporting/2021-01-13_Variable_Dictionary_and_Disclaimer_national_weekly_data.pdf`. I stored this file also locally, because history shows that the ECDC is not very stable in it's accessibility policy.

```{r, echo=TRUE, results='hide'}
## Read data from public url to csv file
data <- readr::read_csv(
  "https://opendata.ecdc.europa.eu/covid19/nationalcasedeath/csv", 
  na = c("", " "))
data


## load vaccination data
data_vac <- readr::read_csv(
  "https://opendata.ecdc.europa.eu/covid19/vaccine_tracker/csv/data.csv", 
  na = c("", " "))
#data_vac

# preprocess and clean data_vac
data_vac <- data_vac %>%
  separate(col = "YearWeekISO", into = c("year", "week"), sep = "-") %>%
  mutate(
    week = str_replace_all(
      string = week, 
      pattern = "W", 
      replacement = "")) %>%
  mutate(
    date =
      convert_week_to_date(x = week, y = year)
  )

# adding a month column 
data_vac <- data_vac %>%
  mutate(date_chr = as.character(date)) %>%
  separate(date_chr, 
           into = c("year2", "month", "day"), 
           sep = "-",
           remove = FALSE) %>%
  mutate(month = as.integer(month)) %>%
  rename(Code = ReportingCountry)


## add country 
#download.file("https://datahub.io/core/country-list/r/data.csv", 
#              destfile = here::here(
#                "data-raw", 
#                "codes_countries"))
data_country_codes <- read_csv(
  "https://datahub.io/core/country-list/r/data.csv")

## join
data_vac <- left_join(data_vac, 
                      data_country_codes) %>%
  rename(country = Name) %>%
  janitor::clean_names()


```

## Convert week number to date
We will do this in steps

 1. Split `year_week` column in `year` and `week`
 1. Convert week numbers to dates

```{r}
convert_week_to_date <- function(x, y){
  map2(
  x, 
  y,
  get_date) %>%
  unlist() %>%
  lubridate::as_date()

}

data_tidy <- data %>%
  separate(col = "year_week", into = c("year", "week")) %>%
  mutate(
    date =
      convert_week_to_date(x = week, y = year)
  )

data_tidy[c(1:4), "date"]

# manual check confirms: week 1 of 2020 was  2019-12-30 to 2020-01-05
# How can the ECDC be so stupid that they do not include a proper date column in their data? 
data_tidy$date %>% unique()


# adding a month column 
data_tidy <- data_tidy %>%
  mutate(date_chr = as.character(date)) %>%
  separate(date_chr, 
           into = c("year2", "month", "day"), 
           sep = "-",
           remove = FALSE) %>%
  mutate(month = as.integer(month))



```

```{r, results='asis'}
country_codes <- str_split(string = params$country_code, pattern = " ") %>% unlist()

countries <- data %>%
  dplyr::filter(country_code %in% country_codes) %>%
  dplyr::select(country) %>%
  unlist() %>%
  unique() %>%
  as.character()

cat("The data is displayed for", glue_collapse(countries, ", ", last = " and "))
```

## Data cleaning
```{r}

selected_dates <- str_split(string = params$from_to_date, pattern = " ") %>% unlist()

date_sequence <- lubridate::as_date(selected_dates)

all_dates <- c(date_sequence[1]:date_sequence[2]) %>% lubridate::as_date()


## Filter for `r params$month` and `r params$geo_id`
## create a conversion table to go from month name to month integer
# convert_months <- tibble(
#   month = c(
#     "january",
#     "february",
#     "march",
#     "april",
#     "may",
#     "june",
#     "july",
#     "august",
#     "september",
#     "october",
#     "november",
#     "december"),
#   month_number = c("01":"12"))
# 
#   month_id <- convert_months %>%
#     dplyr::filter(
#       month %in% m_names$value) 
# id <- c(min(month_id$month_number): max(month_id$month_number))
# 
# ## supplement series
# 
# min(selected_years %>% as.numeric())
# max(selected_years %>% as.numeric())
# 
# 
# 
data_filter <- data_tidy %>%
  dplyr::filter(date %in% all_dates) %>%
#  dplyr::filter(year %in% selected_years) %>%
  dplyr::filter(country_code %in% country_codes)
#   
# char_col <- map_lgl(
#   .x = data_filter,
#   .f = is.numeric 
# )
# 
# country_names <- data_filter$country %>% unique() %>%
#   paste(collapse = ", ")

Caps <- function(x) {
s <- strsplit(x, " ")[[1]]
paste(toupper(substring(s, 1,1)), substring(s, 2),
sep="", collapse=" ")
}

#range <- data_filter$month %>% unique() %>% enframe() %>%
#  arrange(value)

#month_df <- left_join(range, convert_months, by = c("value" = "month_number")) %>%
#  dplyr::mutate(month = firstupper(month))
# 
# min_month <- month_df %>%
#   dplyr::filter(value == min(month_df$value))
# 
# countries_vec <- str_split(
#   country_names, 
#   pattern = ",", 
#   simplify = FALSE
#   ) %>%
#   unlist() %>%
#   trimws() %>%
#   blscrapeR::firstupper()  
#   
# 
# max_month <- month_df %>%
#   dplyr::filter(value == max(month_df$value))

cap_cases <- paste(
  "COVID-19 positive cases from", 
  selected_dates[1], "to", selected_dates[2],
  "for", glue_collapse(countries, ", ", last = " and ")
)
cap_cases

cap_deaths <- paste(
  "COVID-19 related deaths from", 
  selected_dates[1], "to", selected_dates[2],
  "for", glue_collapse(countries, ", ", last = " and ")
)

cap_deaths
data_filter  
```

## Cases
```{r, fig.cap=cap_cases}
#Sys.setlocale("LC_TIME", "English")Sys.setlocale("LC_TIME", "English")Sys.setlocale("LC_TIME", "English")
plot_cases <- data_filter %>%
  na.omit() %>%
  dplyr::filter(indicator == "cases") %>%
  ggplot(aes(
    x = date, 
    y = weekly_count)) +
  geom_point(aes(colour = country)) +
  geom_line(aes(group = country, colour = country)) +
  xlab("Date") +
  ylab("COVID-19 positive tested cases") +
  scale_x_date(breaks="month", date_labels = "%b") +
  xlab(NULL) +
  theme_bw() +
  toolboxr::rotate_axis_labels("x", 90) 

plot_cases
```

## Deaths
```{r, fig.cap=cap_deaths}
## setting month abbreviations to "English"
#Sys.setlocale("LC_TIME", "English")

plot_deaths <- data_filter %>%
  na.omit() %>%
  dplyr::filter(indicator == "deaths") %>%
  ggplot(aes(
    x = date, 
    y = weekly_count)) +
  geom_point(aes(colour = country)) +
  geom_line(aes(group = country, colour = country)) +
  xlab("Date") +
  ylab("COVID-19 positive tested cases") +
  scale_x_date(breaks="month", date_labels = "%b") +
  xlab(NULL) +
  theme_bw() +
  toolboxr::rotate_axis_labels("x", 90) 
plot_deaths
```

## Vaccination - weekly
```{r}

   data_vaccination_total <- data_vac %>%
      dplyr::filter(
        # continent == input$continent,
        country %in% countries) %>%
      group_by(date) %>%
      summarise(total = sum(number_doses_received))
  
data_vac %>%
      dplyr::filter(
       # continent == input$continent,
        country %in% countries) %>%
      group_by(country, date) %>%
      summarise(total = sum(number_doses_received, na.rm = TRUE)) %>%
      ggplot(
        aes(
          x = date,
          y = total
        )
      ) +
      geom_point(aes(colour = country), size = 2) +
      scale_x_date(breaks="month", date_labels = "%b-%y") +
      geom_line(aes(colour = country, group = country), size = 1) +
      ggtitle(
        paste("anti-SARS-CoV2 vaccinations", 
              countries)) +
      ylab("Total count per week") +
      geom_line(
        data = data_vaccination_total, 
        colour = "darkred", 
        size = 2, 
        linetype = "dashed") +
      scale_y_log10() +
    #  toolboxr::rotate_axis_labels("x", 90) +
      theme_bw() -> vaccination_plot
    
vaccination_plot  


ind <- !is.na(data_vac$number_doses_received)

data_vac_no_na <- data_vac[ind, ]

## cummulative

vaccination_plot_cumsum <- data_vac_no_na %>%
  #    na.omit() %>%
      dplyr::filter(
        country %in% countries) %>%
      group_by(country, date) %>%
  summarize(total = sum(number_doses_received, na.rm = TRUE)) %>%
      mutate(total_cummulative = cumsum(total)) %>%
    #  group_by(country, date)
    #  summarize(total = sum(total_cummlative)) %>%
      ggplot(
        aes(
          x = date,
          y = total_cummulative
        )
      ) +
      geom_point(aes(colour = country),size = 2) +
      scale_x_date(breaks="month", date_labels = "%b-%y") +
      geom_line(aes(group = country, colour = country)) +
      ggtitle(
        paste("anti-SARS-CoV2 vaccinations", 
              countries)) +
      ylab("Total count per week") +
      scale_y_log10() +
    #  toolboxr::rotate_axis_labels("x", 90) +
      theme_bw()

vaccination_plot_cumsum


```




## Interactive plot
```{r, fig.height=25, fig.width=18, eval=FALSE}
source(
  here::here(
    "app",
    "shiny",
    "app.R"
  )
)

shinyApp(ui = ui, server = server)
```

